---
title: "Data gathering & cleaning from archival sources"
author: "Rick O. Gilmore"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    self_contained: false
params:
  db_account: rogilmore@psu.edu
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

# Background

This document describes the data gathering and data cleaning steps involved in analyzing some previous optic flow data from the perspective of sex differences.

The original data were estimates of the optic flow speeds and direction patterns derived from head-mounted videos of infants. 
The original video data can be found here: 

Jayaraman, S., Smith, L.B., Raudies, F. & Gilmore, R.O. (2014). Natural Scene Statistics of Visual Experience Across Development and Culture. Databrary. Retrieved April 5, 2019 from http://doi.org/10.17910/B7988V.

This Databrary volume also contains links to talks given at ICIS 2016 and EpiRob 2015.

In addition to the optic flow data, there are data derived from human coders who judged whether the infants were stationary or moving in the videos.

This repo gathers and cleans these data so that we can ask questions about the extent to which sex differences in optic flow experiences seem to occur in our sample.
The cleaned data files will ultimately go back on Databrary so that we can produce a reproducible analysis workflow.

# Data gathering

I'm going to begin with an attempt to categorize the separate datasets (in `data/` by content, focusing on those that appear to be most relevant for the question at hand.

## `pattern-histogram-bins.csv`

```{r}
pat_hist_bins <- read_csv('data/pattern-histogram-bins.csv')
str(pat_hist_bins)
```

This file contains data about the extent to which segments of video (5 s in duration or 150 frames @ 30 frames/s) correlated with one of the optic flow pattern types: left/right, up/down, inward/outward, or clockwise/counter clockwise.
The data report a `Corr.Bin.Ctr` and `Corr.Bin.Count` for each component of a histogram returned for each 5s segment and for each pattern type (`Pattern`).

We also have starting (`Start_fr`) and ending (`End_fr`) video frame values, the `Participant.ID`, age group (`AgeMatchGroup`), participant country of origin (`Country`) to work with.
Finally, each segment was coded by a human observer assigned to judge whether the infant was moving (actively or passively) or stationary during the sampled video segment.

Aside from some challenges in visualizing the data, it should be relatively straightforward to add a `gender` variable to this data file.

## `speed-histogram-bins.csv`

```{r}
speed_hist_bins <- read_csv('data/speed-histogram-bins.csv')
str(speed_hist_bins)
```

This data file contains the results of fitting a gamma ($\Gamma$) distribution to the optic flow speed data for each participant.

According to 

Gilmore, R. O., Raudies, F., & Jayaraman, S. (2015). What accounts for developmental shifts in optic flow sensitivity? 2015 Joint IEEE International Conference on Development and Learning and Epigenetic Robotics (ICDL-EpiRob) (pp. 19–25). Presented at the 2015 Joint IEEE International Conference on Development and Learning and Epigenetic Robotics (ICDL-EpiRob), ieeexplore.ieee.org. Retrieved from http://dx.doi.org/10.1109/DEVLRN.2015.7345450

the following function was fit to the speed histogram:

γ(x;k,θ,a)=axk−1exp(−x/θ)θkΓ(k)

$\gamma(x; \kappa, \theta, \alpha)=\alpha\frac{x^{\kappa - 1}exp(-x^{\theta})}{\theta^{\kappa}\Gamma(\kappa)}$

The data file contains `Participant.ID`, `Country`, `AgeMatchGroup`, and `Motion.Status` variables familiar from `data/pattern-histogram-bins.csv`.

The `DF` and `ChiSquare` variables appear to describe the `goodness-of-fit`.

I think the `Speed` variable refers to each speed bin (in deg/s) by observer motion category {`moving`, `stationary`}.
I will need to reread the paper for more insights.

## `speed-histogram-fits.csv` 

```{r}
speed_hist_fits <- read_csv('data/speed-histogram-fits.csv')
str(speed_hist_fits)
```

These data contain summary statistical measures of the fitted speed distributions: `Mean.spd`, `Median.spd`.
The data also include the fitted parameters from the \Gamma fit, where `a` is the $\alpha$ parameter, `k` is $\kappa$, `theta` is $\theta$, `r2` is the $r^2$ value for the fit.
The `p05`, `p01`, `p005`, and `p001` nominal [0,1] variables may refer to a test between the `stationary` and `moving` conditions.

It should be straightforward to add `gender` and conduct some simple visualizations with these data.

## `wks.sex.csv`

```{r}
wks_sex <- read_csv('data/wks.sex.csv')
str(wks_sex)
```

This file contains easily interpretable participant-specific metadata.

# Data cleaning

## Plan of attack

Participant specific metadata like sex, age, and country of origin are organized within the Databrary spreadsheet.
I will use that source.

It makes the most sense to

- 'strip' these metadata fields from the `data/pattern-histogram-bins.csv`, `data/speed-histogram-fits.csv`, and `data/speed-histogram-bins.csv` files.
- Upload the 'clean' files to an appropriate `materials` folder on Databrary.
- Develop analysis workflows anchored to the Databrary volume for reproducibility.

## Strip redundant fields and resave

```{r}
pattern_histogram_bins <- pat_hist_bins %>%
  select(-Country, -AgeMatchGroup) %>%
  rename(ID = Participant.ID, video = Video, motion_status = Motion.status,
         pattern_type = Pattern, start_fr = Start_fr, end_fr = End_fr,
         n_frames = N_fr,
         corr_bin_ctr = Corr.Bin.Ctr, corr_bin_count = Corr.Bin.Count)
write_csv(pattern_histogram_bins, 'data/pattern_histogram_bins.csv')
```

```{r}
speed_histogram_bins <- speed_hist_bins %>%
  select(-Country, -AgeMatchGroup) %>%
  rename(ID = Participant.ID, motion_status = Motion.status,
         chi_square = ChiSquare, df = DF, p_val = Pval,
         speed_deg_sec = Speed,
         n_obs = N.obs)
write_csv(speed_histogram_bins, 'data/speed_histogram_bins.csv')
```

```{r}
speed_histogram_fits <- speed_hist_fits %>%
  select(-Country, -AgeMatchGroup) %>%
  rename(ID = Participant.ID, motion_status = Motion.status, 
         mean_speed = Mean.spd,
         median_spd = Median.spd, alpha = a, kappa = k)
write_csv(speed_histogram_fits, 'data/speed_histogram_fits.csv')
```

Next, I will upload `pattern_histogram_bins.csv`, `speed_histogram_bins.csv`, and `speed_histogram_fits.csv` to 

Jayaraman, S., Smith, L.B., Raudies, F. & Gilmore, R.O. (2014). Natural Scene Statistics of Visual Experience Across Development and Culture. Databrary. Retrieved April 5, 2019 from http://doi.org/10.17910/B7988V.

These files are in <https://nyu.databrary.org/volume/81/slot/35545/-> or `vol_id = 81`, `session_id = 35545`. 
