---
title: "research_idea_May_10_2019"
author: "Yiming"
date: "5/10/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(tidyverse)
```

# Sex difference in motion perception
### Aims:
      - Aim 1:	It answers whether difference of motion perception is caused by sex (physically inherited) or visual experience (environmental).
      - Aim 2:  It answers whether neural underpinning of sex difference in motion perception.
      - Aim 3:	Investigate whether male and female use different perceptual strategies which cause sex difference

## Preliminary analysis 
Does the speed of optic flow of boy and girl infants differ in their active (walking/crawling) or negative (carrying) visual experience?
It answers whether the activity preference or sex influence motion perception.

### speed comparison
```{r load-data, echo=FALSE}
spd_fits<-read.csv("data/spd_fits.csv", header=TRUE, sep=",")
spd_fits<-spd_fits[!spd_fits$ID == "018MT",]
head(spd_fits)
```
Plot (mean speed as DV)
```{r boxplot-speed, echo=FALSE}
p <- ggplot(spd_fits, aes(x=gender, y=mean_speed, fill = motion_status)) + 
    scale_x_discrete(name = "gender") +
    scale_y_continuous(name = "mean speed") +
    geom_boxplot()+
    theme_bw() +
    theme(plot.title = element_text(size = 14, family = "Tahoma", face = "bold"),
              text = element_text(size = 12, family = "Tahoma"),
              axis.title = element_text(face="bold"),
              axis.text.x=element_text(size = 11)) 
     # + scale_fill_discrete(name="Group",
     #                    breaks=c("T", "U"),
      #                   labels=c("Control", "Down's"))
p

```
```{r lineplot-speed, echo=FALSE}
df<- spd_fits %>%
  group_by(gender, motion_status) %>%
  summarize(total.mean=mean(mean_speed),
            n=n(),
            total.se=sd(mean_speed)/sqrt(n))

p1<-ggplot(df, aes(x=gender, y=total.mean, group=motion_status, color=motion_status))+
         geom_line()+
        labs(x="gender", y="mean speed")  +
        geom_pointrange(aes(ymin=total.mean-total.se, ymax=total.mean+total.se))
         # ggtitle('total translational counts') +
    #  +  scale_colour_hue(name="group",    #Legend label, use darker colors
     #                 breaks=c("T", "U"),
      #              labels=c("Control", "Down's"))+
       
p1
```
### Linear regression
#### 3 IVs: gender, country and motion_status
#### DV: mean_speed
```{r linear-regression-mean, echo=FALSE}
linearMod <- lm(mean_speed ~ gender  + country+ motion_status , data=spd_fits, na.action=na.omit)
summary(linearMod)
```
### Linear Regression
#### DV: median_spd 
```{r linear-regression-median, echo=FALSE}
linearMod <- lm(median_spd ~ gender  + country+ motion_status , data=spd_fits, na.action=na.omit)
summary(linearMod)
```
The insignifant gender difference makes some sense because infants have limited lomomotion ability and are often carried by caregivers.

### pattern comparison
```{r load-patt, echo=FALSE}
patt_hist<-read.csv("data/patt_hist.csv", header=TRUE, sep=",")
patt_hist<-patt_hist[!patt_hist$ID == "018MT",]
head(patt_hist)
```

```{r lineplot, echo=FALSE}
df2<- patt_hist %>%
  group_by(gender, pattern_type,ID) %>%
  summarize(total.mean=mean(corr_bin_count),
            n=n(),
            total.se=sd(corr_bin_count)/sqrt(n))

p1<-ggplot(df2, aes(x=gender, y=total.mean, group=pattern_type, color=pattern_type))+
         geom_line()+
        labs(x="gender", y="frequency of pattern")  +
        geom_pointrange(aes(ymin=total.mean-total.se, ymax=total.mean+total.se))
         # ggtitle('total translational counts') +
    #  +  scale_colour_hue(name="group",    #Legend label, use darker colors
     #                 breaks=c("T", "U"),
      #              labels=c("Control", "Down's"))+
       
p1
```

##	Possible research projects

1.	Does the choice and time for different type of toys (such as cars, dolls) influence the motion perception/spatial ability between boys and girls? Or motion perception influences toy choice? (Aim 1)
   - Age range: 2-7
   - Design: 2 measures: use camera to record children's choice and time on each category of toys with various types of toys (and the time that children makes toys moving); psychophysical experiment on motion perception
    
2.  Do the male infants/children pay more attention to the moving objects (Aim 1)?
   - Age range: 1-3
   - Design:eye-tracking experiment 

3.	Does visual experiences of female and male differ in older age? How is it correlated to motion perception performance/spatial ability? (Aim 1)
   - Age range: 1-18
   - Design: head camera (maybe some data can be collected in PLAY project)
   
4.	Do males have faster response (ERP peak time) and higher activation (ERP/fMRI activation) to motion (Aim 2)?
   - Age range: 1-18
   - Design: ERP (check previous literature) and fMRI (refer to adult study of Murray et al., 2018)
   
5.  Psychological orientation to thing/person (EEG or fMRI; previous finding showed there is same neural activation area and activation to human face)

6.  Does Hormone (such as prenatal androgen) influence the amount of looking at moving objects/male-typed activity interests and engagement? (Aim 1) If so, does hormone correlated with motion perception performance/spatial ability?

7.  (Maybe) When does sex difference occurs? (hard to do in infancy; Aim 1) 



