---
title: "Sex differences in experienced optic flow"
author: "Rick O. Gilmore & Yiming Qian"
date: "`r Sys.time()`"
output:
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    number_sections: true
    code_folding: show
    self_contained: false
params:
  db_account: rogilmore@psu.edu
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Install packages
if (!(require(databraryapi))) {
  devtools::install_github('PLAY-behaviorome/databraryapi')
}
library(databraryapi)

if (!(require(tidyverse))) {
  install.packages('tidyverse')
}
library(tidyverse)
```

# Background

This document analyzes natural optic flow data drawn from 

Jayaraman, S., Smith, L.B., Raudies, F. & Gilmore, R.O. (2014). Natural Scene Statistics of Visual Experience Across Development and Culture. Databrary. Retrieved April 5, 2019 from http://doi.org/10.17910/B7988V.

Infants from India and the U.S. wore a head-mounted camera for several hours during one day.
Jayaraman et al. used computer vision techniques to extract optic flow patterns and speeds from the videos.

Here, we analyze the data to determine whether there is evidence for sex differences in the speeds of experienced optic flow.

# Data gathering

We lightly modified data stored elsewhere and uploaded them to the Jayaraman et al. Databrary volume.

The (CSV format) data files are in a `materials` folder <https://nyu.databrary.org/volume/81/slot/35545/->

## Work plan

We will do the following:

1. Import the CSVs as data frames using the `databraryapi::read_csv_data_as_df()` command from the `databraryapi` package.
The package can be found at <https://github.com/PLAY-behaviorome/databraryapi>.
2. Import the Databrary participant metadata spreadsheet using the `databraryapi::download_session_csv()` command from the `databraryapi` package.
3. Merge the spreadsheet data with the optic flow data files, one by one, using `dplyr` commands.

**Note:**The `setup` chunk should load the `databraryapi` package. If you are not Rick Gilmore, you will need to render this document from an RStudio console as follows: `rmarkdown::render('sex-diffs-in-flow.Rmd', params = list(db_login=<YOUREMAIL@YOURDOMAIN>))`, where you substitute your Databrary account ID (email) for `<YOUREMAIL@YOURDOMAIN>`.

## Import optic flow CSVs

List the CSVs before we import them. But first, let's assign the `vol_id` and `session_id` variables.

```{r list-assets}
vol_id <- 81
session_id = 35545
databraryapi::list_assets_in_session(vol_id, session_id)
```

Isn't reproducible science fun?

Now, we import these files noting the `asset_id` values from the table.
We do this one at a time so we can confirm everything is working.

```{r read-patt-hist}
patt_hist <- databraryapi::read_csv_data_as_df(session_id, asset_id = 165805)
str(patt_hist)
```

```{r read-spd-bins}
spd_bins <- databraryapi::read_csv_data_as_df(session_id, asset_id = 165804)
str(spd_bins)
```

```{r read-spd-fits}
spd_fits <- databraryapi::read_csv_data_as_df(session_id, asset_id = 165803)
str(spd_fits)
```

This is looking good.

## Import participant data

```{r download-session}
databraryapi::login_db(params$db_account)
participants <- databraryapi::download_session_csv(vol_id)
str(participants)
```

We note that the Databrary API is inconsisent in using periods `.` and underscores `_` in field names.
We can clean this before we merge with the other data files and possibly drop some variables that aren't really useful.

# Data cleaning

## Clean `participants`

I want to retain `participant.ID` which will match to `ID`, `participant.gender`, and `context.country`.
It's not essential, but I prefer lower case variable names with underscores as separators.
Oddly, an age variable is not downloaded.
We can address that later.

```{r clean-participants}
participants <- participants %>%
  select(participant.ID, participant.gender, context.country) %>%
  rename(ID = participant.ID, gender = participant.gender, country = context.country)
```

## Merge files

```{r merge-patt-hist}
patt_hist <- dplyr::left_join(patt_hist, participants, by = "ID")
```

It looks like there may be some inconsistencies in our `ID` variable.

### Diagnosing issues with ID

```{r diagnose-id}
unique(participants$ID)
unique(patt_hist$ID)
unique(spd_bins$ID)
unique(spd_fits$ID)
```

It appears that we have optic flow data for $n=$ `r length(unique(spd_fits$ID))` participants, but there are $n=$ `r length(unique(participants$ID))` total in the repository.
This isn't a problem, but I thought we had more optic flow data to analyze.

## Merging the speed data with participants

```{r merge-spd-bins}
spd_bins <- dplyr::left_join(spd_bins, participants, by = "ID")
```

```{r merge-spd-fits}
spd_fits <- dplyr::left_join(spd_fits, participants, by = "ID")
```
# Visualization

## Plan

Yiming, this is for you.
Hi This is Yiming. I would do the following: 
1. Figure out what is spd_bins, patt_hist, spd-fits. Is it the distribution of speed/optic flow pattern in each frame of the video, or the distribution of mean speed/optic flow pattern in all frames of the video, or the distribution of mean speed/optic flow pattern in all frames of the video of all subjects?
2. make plots of speed and optic flow pattern between boys and girls, Indian and American, Stationary and Moving

L


# Analysis
